---
title: "Sham and CACPR Dataset Integration"
author: "Jonathan Nelson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages
```{r load new packages, echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}

if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!requireNamespace('BiocManager', quietly = TRUE)) {install.packages('BiocManager'); require("BiocManager")}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("sctransform")) {install.packages("sctransform"); require("sctransform")}
if (!require("glmGamPoi")) {BiocManager::install('glmGamPoi'); require("glmGamPoi")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("EnhancedVolcano")) {BiocManager::install('EnhancedVolcano'); require("EnhancedVolcano")}
if (!require("DESeq2")) {BiocManager::install('DESeq2'); require("DESeq2")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}
if (!require("car")) {install.packages("car"); require("car")}
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("gghighlight")) {install.packages("gghighlight"); require("gghighlight")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("here")) {install.packages("here"); require("here")}

here::here()

```

# Loading Filtered Seurat Object
These files are post-DoubletFinder.
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}
Sham <- readRDS(here("Outputs", "Sham.rds"))
CACPR <- readRDS(here("Outputs", "CACPR.rds"))



```

# Nuclei Entering the Analysis Pipeline
The following shows the number of nuclei post-QC and post-DoubletFinder. Numbers in `# (10X Count)` are the starting numbers for reference.  
Values appear in the Environment when running the `rmd`.
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}
Sham.nuclei <- nrow(Sham@meta.data) # 7806 (10X Count)
Sham.nuclei
CACPR.nuclei <- nrow(CACPR@meta.data) # 6332 (10X Count)
CACPR.nuclei

```

# Merging Replicates
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

CACPR_Combined <- merge(Sham, y = CACPR, add.cell.ids = c("Sham", "CACPR"), project = "CACPR")

#saveRDS(CACPR_Combined, file = here("Outputs", "CACPR_Combined.rds"))

#To start from this point just reload RDS
#CACPR_Combined <- readRDS("Outputs", "CACPR_Combined.rds")
```

### Sanity Check
```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}

head(colnames(CACPR_Combined))
tail(colnames(CACPR_Combined))
unique(sapply(X = strsplit(colnames(CACPR_Combined), split = "_"), FUN = "[", 1))
table(CACPR_Combined$orig.ident)
head(CACPR_Combined@meta.data)

#To save RAM
rm(Sham, CACPR)
```

## Data Filtering and Integration
```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}

CACPR_Combined

head(CACPR_Combined@meta.data)

VlnPlot(CACPR_Combined, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by = "orig.ident")

VlnPlot(CACPR_Combined, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by = "Condition")

CACPR_Combined <- subset(CACPR_Combined, subset = nFeature_RNA < 5000 & nCount_RNA < 10000)

VlnPlot(CACPR_Combined, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by = "orig.ident")

VlnPlot(CACPR_Combined, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by = "Condition")

CACPR_Combined

CACPR_Combined <- NormalizeData(CACPR_Combined, normalization.method = "LogNormalize", scale.factor = 10000)

CACPR_Combined <- FindVariableFeatures(CACPR_Combined, selection.method = "vst", nfeatures = 2000)

top10 <- head(VariableFeatures(CACPR_Combined), 10)

plot1 <- VariableFeaturePlot(CACPR_Combined)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot2

all.genes <- rownames(CACPR_Combined)
CACPR_Combined <- ScaleData(CACPR_Combined, features = all.genes)

CACPR_Combined <- RunPCA(CACPR_Combined, features = VariableFeatures(object = CACPR_Combined))

print(CACPR_Combined[["pca"]], dims = 1:5, nfeatures = 5)

VizDimLoadings(CACPR_Combined, dims = 1:2, reduction = "pca")

DimPlot(CACPR_Combined, reduction = "pca")

DimHeatmap(CACPR_Combined, dims = 1:2, cells = 500, balanced = TRUE)

ElbowPlot(CACPR_Combined)

CACPR_Combined <- FindNeighbors(CACPR_Combined, dims = 1:15)
CACPR_Combined <- FindClusters(CACPR_Combined, resolution = 0.15)

CACPR_Combined <- RunUMAP(CACPR_Combined, dims = 1:20)

DimPlot(CACPR_Combined, reduction = "umap")
DimPlot(CACPR_Combined, reduction = "umap", group.by = "Condition")

CACPR_Combined.list <- SplitObject(CACPR_Combined, split.by = "Condition")

CACPR_Combined.list <- lapply(X = CACPR_Combined.list, FUN = function(x) {
  x <- NormalizeData(x, verbose = FALSE)
  x <- FindVariableFeatures(x, verbose = FALSE)
})

features <- SelectIntegrationFeatures(object.list = CACPR_Combined.list)
CACPR_Combined.list <- lapply(X = CACPR_Combined.list, FUN = function(x) {
  x <- ScaleData(x, features = features, verbose = FALSE)
  x <- RunPCA(x, features = features, verbose = FALSE)
})

anchors <- FindIntegrationAnchors(object.list = CACPR_Combined.list, reference = c(1, 2), reduction = "rpca", dims = 1:15)
CACPR_Combined <- IntegrateData(anchorset = anchors, dims = 1:15)

CACPR_Combined <- ScaleData(CACPR_Combined, verbose = FALSE)
CACPR_Combined <- RunPCA(CACPR_Combined, verbose = FALSE)
CACPR_Combined <- RunUMAP(CACPR_Combined, dims = 1:15)

CACPR_Combined <- FindNeighbors(CACPR_Combined, reduction = "pca", dims = 1:15)
CACPR_Combined <- FindClusters(CACPR_Combined, resolution = 0.15)

DimPlot(CACPR_Combined, reduction = "umap", label = T)
DimPlot(CACPR_Combined, reduction = "umap", group.by = "Condition")
DimPlot(CACPR_Combined, reduction = "umap", split.by = "Condition")

table(Idents(CACPR_Combined))

table(Idents(CACPR_Combined), CACPR_Combined$Condition)

DefaultAssay(CACPR_Combined) <- "RNA"

FeaturePlot(CACPR_Combined, "Havcr1", split.by = "Condition")

```

## Cluster Synthesis and Naming
```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}

VlnPlot(CACPR_Combined, "nFeature_RNA")
Cluster8 <- FindMarkers(CACPR_Combined, ident.1 = "8")

#Markers from Kirita et al. Figure 1E

markers.to.plot1 <- c("Lrp2",         # Pan PT
                      "Slc5a12",
                      "Slc13a3",
                      "Slc16a9",
                      "Epha7",        # DTL
                      "Akap12",       # DTL
                      "Slc12a1",      # TAL
                      "Slc12a3",      # Pan DCT
                      "Slc8a1",       # DCT2
                      "Slc5a12",      # PT-S1
                      "Scnn1g",       # CNT
                      "Aqp2",         # PC
                      "Kit",          # IC-A
                      "Slc26a4",      # IC-B
                      "Nphs2",        # Pod
                      "Flt1",         # EC
                      "Pdgfrb",       # Mesenchymal
                      "Fyb")          # Immune Macrophages
                            
VlnPlot(CACPR_Combined, 
        features = markers.to.plot1,
        stack = TRUE,
        fill.by = "ident",
        flip = T,
              )

new.cluster.ids.CACPR_Combined <- c("EC", "TAL", "PT", "PT", "Mes", "DCT", "PT", "PC", "PT", "DTL", "IC", "Immune", "Podo", "PT")
names(new.cluster.ids.CACPR_Combined) <- levels(CACPR_Combined)
CACPR_Combined <- RenameIdents(CACPR_Combined, new.cluster.ids.CACPR_Combined)

#Re-level and set active idents
my_levels <- c("PT", "DTL", "TAL", "DCT", "PC", "IC", "Podo", "EC", "Mes", "Immune")
Idents(CACPR_Combined) <- factor(Idents(CACPR_Combined), levels= my_levels)

CACPR_Combined[["CellType"]] <- Idents(object = CACPR_Combined)

#Re-level Sham and CACPR
my_levels <- c("Sham", "CACPR")
CACPR_Combined@meta.data$Condition <- factor(CACPR_Combined@meta.data$Condition, levels= my_levels)

VlnPlot(CACPR_Combined, 
        features = markers.to.plot1,
        stack = TRUE,
        fill.by = "ident",
        flip = T,
              )

DimPlot(CACPR_Combined, reduction = "umap", label = TRUE)

DimPlot(CACPR_Combined, reduction = "umap", label = TRUE, split.by = "Condition")

table(Idents(CACPR_Combined))

table(Idents(CACPR_Combined), CACPR_Combined$Condition)

#saveRDS(CACPR_Combined, here("Temp Outputs", "CACPR_Combined_Named.rds"))

#To start from here just load the RDS
#CACPR_CombinedNamed <- readRDS(here("Temp Outputs", "CACPR_Combined_Named.rds")

FeaturePlot(CACPR_Combined, "Havcr1", split.by = "Condition")

```

# Output Files
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

saveRDS(CACPR_Combined, here("Outputs", "CACPR_Combined_Named.rds"))

#To start from here just load the RDS
#CACPR_Combined <- readRDS(here("Outputs", "CACPR_Combined_Named.rds"))

```

# Session Info
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

sessionInfo()
```
