---
title: "Figure 2"
author: "Jonathan Nelson and Annie Lackey"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---


```{r setup, include = FALSE}
r <- getOption("repos"); r["CRAN"] <- "http://lib.stat.cmu.edu/R/CRAN/"
options(repos = r)

```

```{r load new packages, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("BiocManager")) {install.packages("BiocManager"); require("BiocManager")}
if (!require("tibble")) {install.packages("tibble"); require("tibble")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")} #color
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("DESeq2")) {BiocManager::install('DESeq2'); require("DESeq2")}
if (!require("here")) {install.packages("here"); require("here")}
if (!require("EnhancedVolcano")) {BiocManager::install('EnhancedVolcano'); require("EnhancedVolcano")} # volcano plot
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}


```

# Load Data
```{r load data, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

IRIintegrated <- readRDS(here("Outputs", "IRI_Combined_Named.rds"))
DefaultAssay(IRIintegrated) <- "RNA"
```

# Figure 2A
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

DimPlot(IRIintegrated, reduction = "umap")

```
# Figure 2B
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

markers.to.plot1 <- c("Lrp2",         # Pan PT
                      "Slc12a1",      # TAL
                      "Slc12a3",      # DCT
                      "Aqp2",         # PC
                      "Kit",          # IC
                      "Nphs2",        # Podo
                      "Flt1",         # EC
                      "Pdgfrb",       # Mesenchymal
                      "Fyb")          # Immune Macrophages

DotPlot(IRIintegrated, features = markers.to.plot1, dot.scale = 8, cols = c("dodgerblue2", "coral2")) +  RotatedAxis() + xlab(label = "") + ylab(label = "") 
```
# Figure 2B alt
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

markers.to.plot1 <- c("Lrp2",         # Pan PT
                      "Slc5a12",
                      "Slc13a3",
                      "Slc16a9",
                      "Slc12a1",      # TAL
                      "Slc12a3",      # DCT
                      "Aqp2",         # PC
                      "Kit",          # IC
                      "Nphs2",       # Podo
                      "Flt1",         # EC
                      "Pdgfrb",       # Mesenchymal
                      "Fyb")          # Immune Macrophages

DotPlot(IRIintegrated, features = markers.to.plot1, dot.scale = 8, cols = c("dodgerblue2", "coral2")) +  RotatedAxis() + xlab(label = "") + ylab(label = "") 

```


# Figure 2C
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

DimPlot(IRIintegrated, reduction = "umap", group.by = "Condition")

```
# Figure 2D
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO <-IRIintegrated

t1 <- table(Idents(SO), SO$Condition)[, c("Control", "TwelveHour")]
t1

prop.t1 <- prop.table(t1, margin = 2)
prop.t1

t2 <- as.data.frame(t1)
t2

colnames(t2) <- c('Cell_type', 'Rep', 'Cell_number')
t2

t3 <- t2 %>%
  pivot_wider(names_from = Cell_type, values_from = Cell_number)
t3

t4 <- t3 %>% 
  add_column(Condition = c("Sham", "IRI"))
t4

t5 <- t4 %>%
  pivot_longer(c("PT", "TAL", "DCT", "PC", "IC", "Podo", "EC", "Mes", "Immune"), names_to ="Cell_type", values_to = "Cell_number")
t5

t6 <- t2 %>% 
  pivot_wider(names_from = Cell_type, values_from = Cell_number) %>%
  add_column(Condition = c("Sham", "IRI")) %>%
  pivot_longer(c("PT", "TAL", "DCT", "PC", "IC", "Podo", "EC", "Mes", "Immune"), names_to ="Cell_type", values_to = "Cell_number")

t5 %>%
  arrange(Cell_number) %>%
  mutate(Cell_type = factor(Cell_type, levels=c("PT", "DTL", "TAL", "DCT", "PC", "IC", "Podo", "EC", "Mes", "Immune"))) %>%
ggplot(
       aes(x=Rep,
           y=Cell_number,
           fill=Cell_type)) +     
  geom_bar(position = "fill",
           stat = "identity",
           colour="black") +
  theme_classic() +
  ggtitle("Cell type distribution per sample")

```

# Figure 2E
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

markers.to.plotTreatmentUp.IRI <- c("Plin2", "Pdk4", "Lcn2", "Spp1", "Vmp1", "Ctsd", "Grem2", "Clu", "Gm42031", "Itgav", "Mt1", "Mt2", "Hsp90aa1", "Igfbp3", "Tnc", "Abi1", "Rgs1", "S100a11")

DotPlot(IRIintegrated, features = markers.to.plotTreatmentUp.IRI, dot.scale = 8, cols = c("dodgerblue2", "coral2"), split.by = "Condition") +  RotatedAxis() + xlab(label = "") + ylab(label = "IRI Enriched") 

```


# Figure 2E
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

markers.to.plotTreatmentDown.IRI <- c("Sugct", "Hnf4aos", "Egf", "Abca13", "Klhl3", "Wnk1", "Rhcg", "Apela", "Tmem117", "Slc16a7", "Gm26917", "Opcml", "Samd12", "Chrm3", "Abca8a", "Stim1", "Rbpj", "Fut9")

DotPlot(IRIintegrated, features = markers.to.plotTreatmentDown.IRI, dot.scale = 8, cols = c("dodgerblue2", "coral2"), split.by = "Condition") +  RotatedAxis() + xlab(label = "") + ylab(label = "IRI Depleted") 

```






markers.to.plotTreatmentUp.IRI <- c("Neat1", "Col18a1", "Spp1", "Cdh9", "Rtn4", "Cep85", "Adk", "Efna5", "Pkhd1", "Atxn1", "Pde10a", "Sdc4", "Malat1", "Fkbp5", "Abi1", "Tnc", "Mt1", "Fth1")

markers.to.plotTreatmentDown.IRI <- c("Slc34a1", "Keg1","Egf", "Slc12a1", "Wnk1", "Abca13", "Col4a3", "Ranbp3l", "Tmem117", "Pde4b", "Pcthd4", "H2-Q7", "Rora", "Tmtc1", "Lbd2", "Arhgap24", "H2-K1", "Cd74" )

DotPlot(IRIintegrated, features = markers.to.plotTreatmentUp.IRI, dot.scale = 8, cols = c("dodgerblue2", "coral2"), split.by = "Condition") +  RotatedAxis() + xlab(label = "") + ylab(label = "IRI Enriched") 


markers.to.plotTreatmentDown.IRI <- c("Sugct", "Hnf4aos", "Egf", "Abca13", "Klhl3", "Wnk1", "Rhcg", "Apela", "Tmem117", "Slc16a7", "Gm26917", "Opcml", "Samd12", "Chrm3", "Abca8a", "Stim1", "Rbpj", "Fut9")

DotPlot(IRIintegrated, features = markers.to.plotTreatmentDown.IRI, dot.scale = 8, cols = c("dodgerblue2", "coral2"), split.by = "Condition") +  RotatedAxis() + xlab(label = "") + ylab(label = "IRI Depleted") 



```



#IRI
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


DimPlot(IRIintegrated, reduction = "umap", label = T)
DimPlot(IRIintegrated, reduction = "umap", group.by = "treatment")
DimPlot(IRIintegrated, reduction = "umap", split.by = "treatment")

markers.to.plot1 <- c("Lrp2",         # Pan PT
                      "Slc5a12",
                      "Slc13a3",
                      "Slc16a9",
                      "Epha7",        # DTL
                      "Slc12a1",      # TAL
                      "Slc12a3",      # Pan DCT
                      "Slc8a1",       # DCT2
                      "Scnn1g",       # CNT
                      "Aqp2",         # PC
                      "Kit",          # IC-A
                      "Slc26a4",      # IC-B
                      "Nphs2",        # Pod
                      "Flt1",         # EC
                      "Pdgfrb",       # Mesenchymal
                      "Fyb")          # Immune Macrophages

VlnPlot(IRIintegrated, 
        features = markers.to.plot1,
        stack = TRUE,
        fill.by = "ident",
        flip = T,
              )

DotPlot(IRIintegrated, features = markers.to.plot1, dot.scale = 8, cols = c("dodgerblue2", "coral2")) +  RotatedAxis() + xlab(label = "") + ylab(label = "") 

SO <-IRIintegrated

t1 <- table(Idents(SO), SO$treatment)[, c("Control", "TwelveHour")]
t1

prop.t1 <- prop.table(t1, margin = 2)
prop.t1

t2 <- as.data.frame(t1)
t2

colnames(t2) <- c('Cell_type', 'Rep', 'Cell_number')
t2

t3 <- t2 %>%
  pivot_wider(names_from = Cell_type, values_from = Cell_number)
t3

t4 <- t3 %>% 
  add_column(Condition = c("Control", "TwelveHour"))
t4

t5 <- t4 %>%
  pivot_longer(c("PT", "TAL", "DCT", "PC", "IC", "Podo", "EC", "Mes", "Immune"), names_to ="Cell_type", values_to = "Cell_number")
t5

t6 <- t2 %>% 
  pivot_wider(names_from = Cell_type, values_from = Cell_number) %>%
  add_column(Condition = c("Control", "TwelveHour")) %>%
  pivot_longer(c("PT", "TAL", "DCT", "PC", "IC", "Podo", "EC", "Mes", "Immune"), names_to ="Cell_type", values_to = "Cell_number")

t5 %>%
  arrange(Cell_number) %>%
  mutate(Cell_type = factor(Cell_type, levels=c("PT", "TAL", "DCT", "PC", "IC", "Podo", "EC", "Mes", "Immune"))) %>%
ggplot(
       aes(x=Rep,
           y=Cell_number,
           fill=Cell_type)) +     
  geom_bar(position = "fill",
           stat = "identity",
           colour="black") +
  theme_classic() +
  ggtitle("Cell type distribution per sample")

DotPlot(IRIintegrated, features = markers.to.plotTreatmentUp.IRI, dot.scale = 8, cols = c("dodgerblue2", "coral2"), split.by = "treatment") +  RotatedAxis() + xlab(label = "") + ylab(label = "IRI Enriched") 

DotPlot(IRIintegrated, features = markers.to.plotTreatmentDown.IRI, dot.scale = 8, cols = c("dodgerblue2", "coral2"), split.by = "treatment") +  RotatedAxis() + xlab(label = "") + ylab(label = "IRI Depleted") 

DotPlot(IRIintegrated, features = markers.to.plotTreatmentUp.IRI, dot.scale = 8, cols = c("dodgerblue2", "coral2"), split.by = "treatment") +  RotatedAxis() + xlab(label = "") + ylab(label = "IRI Enriched") 

DotPlot(IRIintegrated, features = markers.to.plotTreatmentDown.IRI, dot.scale = 8, cols = c("dodgerblue2", "coral2"), split.by = "treatment") +  RotatedAxis() + xlab(label = "") + ylab(label = "IRI Depleted") 


```

#Highlight PT cluster

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


DimPlot(IRIintegrated, split.by= "Condition" , cells.highlight=WhichCells(IRIintegrated, idents = c("PT")), cols.highlight = c("#FF575B"), cols= "grey")

DimPlot(IRIintegrated, cells.highlight=WhichCells(IRIintegrated, idents = c("PT")), cols.highlight = c("#FF575B"), cols= "grey") +
  ggtitle("IRI") +
  theme(plot.title = element_text(hjust = 0.5))



DimPlot(IRIintegrated, split.by="treatment", cells.highlight=WhichCells(IRIintegrated, idents = c("PT")), cols.highlight = c("#FF575B"), cols= "grey")

DimPlot(IRIintegrated, cells.highlight=WhichCells(IRIintegrated, idents = c("PT")), cols.highlight = c("#FF575B"), cols= "grey")+
  ggtitle("IRI") +
  theme(plot.title = element_text(hjust = 0.5))

VlnPlot(IRIintegrated, "Havcr1", split.by =  "Condition")

VlnPlot(IRIintegrated, "Havcr1", split.by =  "treatment")

VlnPlot(IRIintegrated, "Slc34a1", split.by =  "Condition")

VlnPlot(IRIintegrated, "Slc34a1", split.by =  "treatment")

VlnPlot(IRIintegrated, "Ehhadh", split.by =  "Condition")

VlnPlot(IRIintegrated, "Ehhadh", split.by =  "treatment")




```



## Session Info

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
sessionInfo()
```



