---
title: "IRI and CACPR Cluster Defining DEGs"
author: "Jonathan Nelson and Annie Lackey"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load new packages, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("BiocManager")) {install.packages("BiocManager"); require("BiocManager")}
if (!require("tibble")) {install.packages("tibble"); require("tibble")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")} #color
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("DESeq2")) {BiocManager::install('DESeq2'); require("DESeq2")}
if (!require("here")) {install.packages("here"); require("here")}
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if (!require("car")) {install.packages("car"); require("car")}
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}


```


# Load CACPR Dataset
```{r load data, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

CACPRintegrated <- readRDS(here("Outputs", "CACPR_Combined_Named.rds"))
DefaultAssay(CACPRintegrated) <- "RNA"
DimPlot(CACPRintegrated, reduction = "umap")

table(Idents(CACPRintegrated))
prop.table(table(Idents(CACPRintegrated)))

```


# Subset PT and Re-Cluster
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

head(CACPRintegrated@meta.data)

CACPRintegrated_PT <- subset(CACPRintegrated, CellType == "PT")

DimPlot(CACPRintegrated_PT, reduction = "umap")

CACPRintegrated_PT <- FindVariableFeatures(CACPRintegrated_PT, selection.method = "vst", nfeatures = 2000)

top10 <- head(VariableFeatures(CACPRintegrated_PT), 10)

plot1 <- VariableFeaturePlot(CACPRintegrated_PT)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot2

all.genes <- rownames(CACPRintegrated_PT)
CACPRintegrated_PT <- ScaleData(CACPRintegrated_PT, features = all.genes)

CACPRintegrated_PT <- RunPCA(CACPRintegrated_PT, features = VariableFeatures(object = CACPRintegrated_PT))

print(CACPRintegrated_PT[["pca"]], dims = 1:5, nfeatures = 5)

VizDimLoadings(CACPRintegrated_PT, dims = 1:2, reduction = "pca")

DimPlot(CACPRintegrated_PT, reduction = "pca")

DimHeatmap(CACPRintegrated_PT, dims = 1:2, cells = 500, balanced = TRUE)

ElbowPlot(CACPRintegrated_PT)

CACPRintegrated_PT <- FindNeighbors(CACPRintegrated_PT, dims = 1:10)
CACPRintegrated_PT <- FindClusters(CACPRintegrated_PT, resolution = 0.15)

CACPRintegrated_PT <- RunUMAP(CACPRintegrated_PT, dims = 1:20)

DimPlot(CACPRintegrated_PT, reduction = "umap")
DimPlot(CACPRintegrated_PT, reduction = "umap", group.by = "Condition")

CACPRintegrated_PT.list <- SplitObject(CACPRintegrated_PT, split.by = "Condition")

CACPRintegrated_PT.list <- lapply(X = CACPRintegrated_PT.list, FUN = function(x) {
  x <- NormalizeData(x, verbose = FALSE)
  x <- FindVariableFeatures(x, verbose = FALSE)
})

features <- SelectIntegrationFeatures(object.list = CACPRintegrated_PT.list)
CACPRintegrated_PT.list <- lapply(X = CACPRintegrated_PT.list, FUN = function(x) {
  x <- ScaleData(x, features = features, verbose = FALSE)
  x <- RunPCA(x, features = features, verbose = FALSE)
})

anchors <- FindIntegrationAnchors(object.list = CACPRintegrated_PT.list, reference = c(1, 2), reduction = "rpca", dims = 1:10)
CACPRintegrated_PT <- IntegrateData(anchorset = anchors, dims = 1:10)

CACPRintegrated_PT <- ScaleData(CACPRintegrated_PT, verbose = FALSE)
CACPRintegrated_PT <- RunPCA(CACPRintegrated_PT, verbose = FALSE)
CACPRintegrated_PT <- RunUMAP(CACPRintegrated_PT, dims = 1:10)

CACPRintegrated_PT <- FindNeighbors(CACPRintegrated_PT, reduction = "pca", dims = 1:10)
CACPRintegrated_PT <- FindClusters(CACPRintegrated_PT, resolution = 0.15)

DimPlot(CACPRintegrated_PT, reduction = "umap", label = T)
DimPlot(CACPRintegrated_PT, reduction = "umap", group.by = "Condition")
DimPlot(CACPRintegrated_PT, reduction = "umap", split.by = "Condition")

table(Idents(CACPRintegrated_PT))

table(Idents(CACPRintegrated_PT), CACPRintegrated_PT$Condition)

DefaultAssay(CACPRintegrated_PT) <- "RNA"

FeaturePlot(CACPRintegrated_PT, "Havcr1", split.by = "Condition")

DimPlot(CACPRintegrated_PT, reduction = "umap", label = T)
DimPlot(CACPRintegrated_PT, reduction = "umap", group.by = "Condition")


markers.to.plot1 <- c("Lrp2",         # Pan PT
                      "Slc5a12",
                      "Slc13a3",
                      "Slc16a9",
                      "percent.mt")
        
                            
VlnPlot(CACPRintegrated_PT, 
        features = markers.to.plot1,
        stack = TRUE,
        fill.by = "ident",
        flip = T,
              )

table(Idents(CACPRintegrated_PT), CACPRintegrated_PT$Condition)
prop.table(table(Idents(CACPRintegrated_PT), CACPRintegrated_PT$Condition))


```





CellTypeList <- unlist(levels(CACPRintegrated@meta.data$CellType))

for(i in 1:length(CellTypeList)){
  
DEGs <- FindMarkers(CACPRintegrated, ident.1 = CellTypeList[i])

DEGs <- DEGs %>%
  filter(p_val_adj < 0.05) %>%
  rownames_to_column(var="gene")

DEGs = DEGs[order(DEGs[,"avg_log2FC"], decreasing = TRUE),]

assign(paste("CACPRintegrated", CellTypeList[i], sep = "."), DEGs, envir = globalenv())

print(paste("CACPR", CellTypeList[i], "Up"))
top25 <- head(DEGs, 25)
print(top25)

top3 <- head(DEGs$gene, 3)

for (j in top3) {
  B <- VlnPlot(CACPRintegrated, j) + labs(title= j) + xlab(element_blank())
  print(B)
  }

}

#remove objects so that entire environment is only DEG lists to save

rm(DEGs, CACPRintegrated, i, CellTypeList, top3, top25, j, B)

save.image(file = here("GEO", "Outputs", "CACPR Cluster Defining DEGs.RData"))

allobj <- ls()
alldf <- sapply(allobj, get, simplify = FALSE)
alldf <- alldf[sapply(alldf, inherits, "data.frame")]

write.xlsx(alldf, sheetName = names(alldf), file = here("GEO", "Outputs", "CACPR Cluster Defining DEGs.xlsx"))

rm(list = ls())
```


# Load IRI Dataset
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

IRIintegrated <- readRDS(here("GEO", "Outputs", "IRI_Combined_Named.rds"))
DefaultAssay(IRIintegrated) <- "RNA"
DimPlot(IRIintegrated, reduction = "umap")

table(Idents(IRIintegrated))
prop.table(table(Idents(IRIintegrated)))
        
```


# DEG Markers for IRI DAtaset
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

CellTypeList <- unlist(levels(IRIintegrated@meta.data$CellType))

for(i in 1:length(CellTypeList)){

DEGs <- FindMarkers(IRIintegrated, ident.1 = CellTypeList[i])  

DEGs <- DEGs %>%
  filter(p_val_adj < 0.05) %>%
  rownames_to_column(var="gene")

DEGs = DEGs[order(DEGs[,"avg_log2FC"], decreasing = TRUE),]

assign(paste("IRIintegrated", CellTypeList[i], sep = "."), DEGs, envir = globalenv())

print(paste("IRI", CellTypeList[i], "Up"))

print(paste("IRIintegrated", CellTypeList[i], "Up"))
top25 <- head(DEGs, 25)
print(top25)

top3 <- head(DEGs$gene, 3)

for (j in top3) {
  B <- VlnPlot(IRIintegrated, j) + labs(title= j) + xlab(element_blank())
  print(B)
}

assign(paste("IRIintegrated", CellTypeList[i], sep = "."), DEGs, envir = globalenv())

}

#remove objects so that entire environment is only DEG lists to save
rm(DEGs, IRIintegrated, i, CellTypeList, top3, top25, j, B)

save.image(file = here("GEO", "Outputs", "IRI Cluster Defining DEGs.RData"))

allobj <- ls()
alldf <- sapply(allobj, get, simplify = FALSE)
alldf <- alldf[sapply(alldf, inherits, "data.frame")]

write.xlsx(alldf, sheetName = names(alldf), file = here("GEO", "Outputs", "IRI Cluster Defining DEGs.xlsx"))

```

## Session Info

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
sessionInfo()
```


  